Co-citations in context

### 2019-05-16 cocit_r.tex
trimmed off extraneous LaTeX

\section{Introduction}
Citation and network analysis of  scientific literature reveals information on semantic relationships between publications, collaboration between scientists, and the practice of citation itself~\citep{de_solla_price_networks_1965,newman_structure_2001,Shi:2010:CHI:1816123.1816131,patience_pmid28560354}. Co-citation, the frequency with which two documents are cited together in other documents~\citep{small_co-citation_1973,marshakova-shaikevich_co-citation_1973}, provides additional insights, including the identification of semantically related documents and fields. Co-citations, bibliographic coupling and direct citation have been compared for XX.

In an intriguing study, co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS) was used to document novel combinations of prior research and to characterize highly cited papers~\citep{uzzi_atypical_2013}. In this study, $nC2$ combinations of the references in each article were generated and then the reference identifiers were replaced with journal identifiers to generate a journal-pair. Observed frequencies of journal pairs across the dataset were then normalized to expected values generated by a random graph model using a \emph{z-score} (Materials and Methods). Thus, every article was associated with multiple z-scores. For each article, positional statistics of these z-scores were calculated to describe conventionality (high conventionality (HC) if the median z-score for an article was greater than the median z-score of all articles and low conventionality for the inverse (LC).) Similarly with novelty (high novelty (HN) if the tenth percentile of z-scores for an article was less than zero and low novelty (LN) for the inverse). Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCLN, with all four combinations being possible. The finding of interest was that HCHN articles were twice as likely to be found in highly cited papers (defined as the top 10\% of all papers published in a given year considering the count of citations accumulated in the first 8 years since publication).

Key to the study is the random graph approach, where every reference is randomly replaced by another article published in the same year as the cited reference to generate a null model for comparison. However, the approach inadequately accounts for observed citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017} since (i) it treats potential substitutions as equi-probable, irrespective of disciplinary origin (ii) does not consider the frequency with which references are cited in the dataset. In illustration, a reference from musicology can replace one in quantum physics during the simulation process. Similarly, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once. Accordingly, the expected value calculations used in normalizing observed journal-pair-frequencies can be reasonably questioned on grounds of model mis-specification.

A follow-up study using a smaller dataset of 12 million articles, also suggests that Uzzi and colleagues do not fully account for disciplinary and journal effects~\citep{boyack_vs_uzzi_2014}. While the two studies are conceptually similar, the authors of the latter study used Scopus data and the K50 measurement, a derivative of the $\chi^2$ method, for normalizing observed to expected journal-pair frequencies. Boyack and Klavans note that 'only 64.4\%  of  243  WoS  subject  categories` in the Uzzi study met the criterion of having the highest probability of hit papers in the HCHN category. Further they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, that three multidisciplinary journals accounted for 9.4\% of all atypical combinations. suggesting strong effects from both disciplines and journals.

Despite different methods used to generate expected values, both Uzzi and Boyack measured observed and expected journal-pair frequencies across disciplines and then examined subsets of publications for disciplinary effects. Since disciplinary origin was not considered when generating expected values in their studies, we hypothesized that analyzing semantically related sets of documents instead of all publications in a literature database would create disciplinary networks. From these, expected values could be calculated, with the effect of reducing model misspecification and better simulating citation practice, in turn leading to different and more relevant conclusions.  We also developed an efficient algorithm that enabled us to conduct a greater number of simulations in a reasonable time. Thus, while Uzzi et al. averaged the results of 10 simulations for each year in their dataset, we performed 1,000 simulations to ensure greater coverage of observed journal pairs. Our study on three semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion that HCHN articles have a greater probability of being represented in highly cited publications.

@@

\newpage
\section{Sample citations}
For general information on the correct form for citations using
the APA 6 format, see the following sites:
\href{https://owl.english.purdue.edu/owl/resource/560/02/}
{APA 6, In-text citations, The Basics} and
\href{https://owl.english.purdue.edu/owl/resource/560/03/}
{APA 6, In-text citations}

\section{Natbib citation mark up}

\subsection{Single citations}
\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90}+&\dogray Jones et al. (1990)\\
\verb+\citet[chap. 2]{jon90}+&\dogray Jones et al. (1990, chap. 2)\\
    \verb+\citep{jon90}+	    &\dogray   	(Jones et al., 1990)\\
    \verb+\citep[chap. 2]{jon90}+ 	&\dogray    	(Jones et al., 1990, chap. 2)\\
    \verb+\citep[see][]{jon90}+ 	 &\dogray    	(see Jones et al., 1990)\\
    \verb+\citep[see][chap. 2]{jon90}+ 	&\dogray    	(see Jones et al., 1990, chap. 2)\\
    \verb+\citet*{jon90}+ 	    &\dogray    	Jones, Baker, and Williams (1990)\\
    \verb+\citep*{jon90}+	    & \dogray   	(Jones, Baker, and Williams,
    1990) \\
\end{tabular}

For example, some citations from the NETNbibsamp.bib database:

citet: \citet{bullmore2009complex}, citep: \citep{gomez2009analysis}, and
citep*: \citep*{de2012cortical}

\subsection{Multiple citations}
Multiple citations may be made by including more than one citation
key in the \verb+\cite+ command argument.

\noindent
\begin{tabular}{ll}
\bf Type&\bf Results\\
\hline
\verb+\citet{jon90,jam91}+&\dogray Jones et al. (1990); James et al. (1991)\\
\verb+\citep{jon90,jam91}+&\dogray (Jones et al., 1990; James et al. 1991)\\
\verb+\citep{jon90,jon91}+&\dogray (Jones et al., 1990, 1991)\\
\verb+\citep{jon90a,jon90b}+&\dogray (Jones et al., 1990a,b)\\
\end{tabular}

For example, multiple citations from the bibsamp.bib database:
citet: \citet{nooner2012nki,hutchison2013resting},
 citep:
\citep{tagliazucchi2012dynamic, de2012cortical}

As you see, the citations are automatically hyperlinked to their
reference in the bibliography.

\newpage

@@

\acknowledgments
Enter your acknowledgments here.

%% ie.,

% The authors thank Laurence Aitchison for fruitful discussions.  RKN
% and PD received funding from the Gatsby Charitable Foundation. Y-AB,
% RBS, KC and PS received funding from Canadian Institutes of Health
% Research grant $MOP74577$,
% Fond de recherche Qu\'{e}bec - Sant\'{e} (Group grant to the Groupe 
% de recherche en neurobiologie comportementale, Shimon Amir, P.I.), and
% Concordia University Research Chair (Tier I).

\authorcontributions 
Who helped formulate the project, who supplied data, analyses and
experiments, etc.

%% ie.
%% Project was formulated by RKN, PD, PS,
%% based on substantial data, analyses and experiments of Y-AB, KC, RS,
%% PS. RKN, PD formalised the model, RKN implemented and ran the model;
%% RKN analysed the molecular ethogram data; Y-AB formalised and
%% implemented a CTMC model. All authors wrote the manuscript.

\section{Making Your Bibliography for a Network Neuroscience Article}
{\it Network Neuroscience} uses the APA author-date  bibliography style,
apacite.bst. For more
information on apacite, for examples in how to make your .bib file and more, see:\\
\href{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}
{http://mirror.jmu.edu/pub/CTAN/biblio/bibtex/contrib/apacite/apacite.pdf}

\noindent
(In spite of the mention of apacite cite commands, please use only
Natbib commands for in text citations, as shown above.)

\subsection{BibTeX}
You will need to use BibTeX to form your bibliography; typing in the
references would be
a huge and unpleasant task. Look at the NETNSample.bbl file and you'll see why
typing in the bibitems would be difficult.

For a good basic introduction to using BibTeX, see\\
\href{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}
{https://www.economics.utoronto.ca/osborne/latex/BIBTEX.HTM}

When you use BibTeX, the form of the bibliography will be correct. You
don't need to supply a bibliography style, since that is built into
the stjour.cls file when the NETN option is used
(\verb+\documentclass[NETN]{stjour}+).

\newpage
\subsection{Sample citations}

Here are some samples using \verb+\citep{}+:\\
\citep{bullmore2009complex,
gomez2009analysis,
rubinov2011weight,
power2014methods,
sporns2015modular,
scheeringa2012eeg,
fortunato2007resolution,
reichardt2006statistical,
smith2009correspondence,
sporns2011networks,
liegeois2014cerebral}

And more using \verb+\citet{}+\\
\citet{allen2012tracking,
calhoun2014chronnectome,
liu2013time,
fisher1915frequency,
gonzalez2014spatial,
damaraju2014dynamic,
hutchison2013resting,
de2012cortical,
tagliazucchi2012dynamic,
nooner2012nki}

>>>>

\section{Introduction}
Citation and network analysis of  scientific literature reveals information on semantic relationships between publications, collaboration between scientists, and the practice of citation itself~\citep{garfield_citation_1955,de_solla_price_networks_1965,newman_structure_2001,Shi:2010:CHI:1816123.1816131,patience_pmid28560354}. Co-citation, the frequency with which two documents are cited together in other documents~\citep{small_co-citation_1973,marshakova-shaikevich_co-citation_1973}, provides additional insights, including the identification of semantically related documents and fields. Co-citations, bibliographic coupling and direct citation have been compared for XX.

In an intriguing study, co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS) was used to document novel combinations of prior research and to characterize highly cited papers~\citep{uzzi_atypical_2013}. In this study, pairwise combinations of the references in each article were generated to create a multi-set of journal-pairs. Observed frequencies of these journal pairs across the dataset were computed and normalized to expected values generated by Monte Carlo simulations from a random graph model. Normalized, journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores. For each article, positional statistics of these z-scores were calculated to describe conventionality (high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the inverse (LC).) Similarly with novelty (high novelty (HN)- if the tenth percentile of z-scores for an article was less than zero and low novelty (LN) for the inverse). Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCLN, with all four combinations being possible. The finding of considerable interest was that HCHN articles were twice as likely to be found in highly cited papers (defined as the top 10\% of all papers published in a given year considering the count of citations accumulated in the first 8 years since publication).

Key to the Uzzi study, however, is the random graph approach, where every reference is randomly replaced by another article published in the same year as the reference to generate a null model for comparison. However, the approach inadequately accounts for observed citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017} since (i) it treats potential substitutions as equi-probable, irrespective of disciplinary origin (ii) does not consider the frequency with which references are cited in the dataset. In illustration, a reference from musicology can replace one in quantum physics during the simulation process. Similarly, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once. Accordingly, the expected value calculations used in normalizing observed journal-pair-frequencies can be reasonably questioned on grounds of model mis-specification.

A follow-up study using a smaller dataset of 12 million articles, also suggests that Uzzi and colleagues do not fully account for disciplinary and journal effects~\citep{boyack_vs_uzzi_2014}. While the two studies are conceptually similar, the authors of the latter study used Scopus data and the K50 measurement, a derivative of the $\chi^2$ method, rather than Monte Carlo simulations for normalizing observed to expected journal-pair frequencies. Boyack and Klavans note that `only 64.4\%  of  243  WoS  subject  categories' in the Uzzi study met the criterion of having the highest probability of hit papers in the HCHN category. Further, they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, that three multidisciplinary journals accounted for 9.4\% of all atypical combinations, suggesting strong effects from both disciplines and journals.

Despite different methods used to generate expected values, both Uzzi and Boyack measured observed and expected journal-pair frequencies across disciplines and then examined subsets of publications for disciplinary effects. Since disciplinary origin was not considered when generating expected values in their studies, we chose instead to analyze semantically related sets of documents (disciplinary networks) instead of all publications in a bibliographic database. We used keyword searches of the scientific literature to create three citation networks themed around broad search terms. For each of these, we calculated expected values using a refined random graph model that accounted for existing citation frequencies in these networks and an efficient simulation algorithm that randomly permuted references grouped by year of publication. Thus, while Uzzi et al. averaged the results of 10 simulations for each year in their dataset, we averaged 1,000 simulations to ensure greater coverage of rarely observed journal pairs We hypothesized that our approach would  reduce model misspecification and better simulate citation practice, in turn leading to different and more relevant conclusions.  Our study on these semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion of universality in HCHN articles have a greater probability of being represented in highly cited publications.

\section{Materials and Methods}
\section{Results}
\section{Discussion}

@@

@@

\acknowledgments
We deeply thank Kevin Boyack and Dick Klavans for critically constructive discussions. We thank the authors of Uzzi et al. (2013) for sharing their Python simulation code. Research and development reported in this publication was partially supported by Federal funds from the National Institute on Drug Abuse, National Institutes of Health, US Department of Health and Human Services, under Contract Nos. HHSN271201700053C (N43DA-17-1216) and HHSN271201800040C (N44DA-18-1216). The content of this publication is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health. TW receives funding from the Grainger Foundation. All the code used in this study is freely available from a Github site. Access to bibliographic data requires a license from Clarivate Analytics, which had no role in 
funding, experimental design, review of results, and conclusions presented.

\authorcontributions 
This study was designed by GD, JB, SD, and TW. Simulations and analysis were performed by AD, GC, JB, and SD. Infrastructure, and workflows used to generate data used in this study were developed by AD, DK, SL, SD, and GC.  All authors reviewed and commented on the manuscript, which was written by GC, JB, and TW.
